---
title: "Comparing mlr3pipelines to mlr and Python"
author: "Florian Pfisterer"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Introduction to mlr3pipelines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>"
)
set.seed(8008135)
compiler::enableJIT(0)
library("mlr3")
library("mlr3pipelines")
```


# Comparing mlr3pipelines to other frameworks

## mlr3pipelines vs. mlr

In order to showcase the benefits of **mlr3pipelines** over **mlr's** `Wrapper` mechanism,
we compare the case of imputing missing values before filtering the top 2 features and then applying a learner.

While **mlr** wrappers are generally less verbose and require a little fewer code, this heavily inhibits flexibility.
Wrappers can generally not process data in parallel, i.e.
### mlr

```{r, eval = FALSE}
library(mlr)
# We first create a learner
lrn = makeLearner("classif.rpart")
# Wrap this learner in a FilterWrapper
lrn.wrp = makeFilterWrapper(lrn, fw.abs = 2L)
# And wrap the resulting wrapped learner into an ImputeWrapper.
lrn.wrp = makeImputeWrapper(lrn.wrp)

# Afterwards, we can train the resulting learner on a task
train(lrn, iris.task)
```

### mlr3pipelines

```{r, eval = FALSE}
library(mlr3)
library(mlr3pipelines)
library(mlr3featsel)

impute = PipeOpImpute$new()
filter = PipeOpFilter$new(filter = FilterVariance$new(), param_vals = list(filter.nfeat = 2L))
rpart = PipeOpLearner$new(mlr_learners$get("classif.rpart"))

# Assemble the Pipeline
pipeline = impute %>>% filter %>>% rpart
# And convert to a 'GraphLearner'
learner = GraphLearner$new(id = "Pipeline", pipeline)
# We want to filter out all but 2 features

e = Experiment$new(task = mlr_tasks$get("iris"), learner = learner)
e$train()
```

The fact that **mlr's** wrappers have to be applied inside-out, i.e. in the reverse order
is often confusing. This is way more straight-forward in `mlr3pipelines`, where
we simply chaing the different methods using `%>>%`. Additionally, `mlr3pipelines`
offers way greater possibilities with respect to the kinds of Pipelines that can be constructed.
In `mlr3pipelines`, we allow for the construction of parallel and conditional pipelines.
This was previously not possible.


## mlr3pipelines vs. sklearn.pipeline.Pipeline

In order to broaden the horizon, we compare to **Python** **sklearn's** `Pipeline` methods.
`sklearn.pipeline.Pipeline` sequentially applies a list of transforms before fitting a final estimator.
Intermediate steps of the pipeline are `transforms`, i.e. steps that can learn from the data,
but also transform the data while it flows through it.
The purpose of the pipeline is to assemble several steps that can be cross-validated together while setting different parameters.
For this, it enables setting parameters of the various steps

It is thus conceptually very similar to `mlr3pipelines`.
Similarly to `mlr3pipelines`, we can tune over a full `Pipeline` using various tuning methods.
`Pipeline` mainly supports linear pipelines. This means, that it can execute parallel steps, such as
for example **Bagging**, but it does not support conditional execution, i.e. `PipeOpBranch`.

At the same time, the different `transforms` in the pipeline can be cached, which makes tuning over the configuration
space of a `Pipeline` more efficient, as executing some steps multiple times can be avoided.

We compare functionality available in both `mlr3pipelines` and `sklearn.pipeline.Pipeline` to give a
comparison.


### Tuning Feature Selection and PCA before applying a SVM

The following example obtained from the **sklearn** documentation showcases a **Pipeline**
that first Selects a feature and performs PCA on the original data, concatenates the resulting datasets
and applies a Support Vector Machine.


### sklearn

```{python, eval = FALSE}
from sklearn.pipeline import Pipeline, FeatureUnion
from sklearn.model_selection import GridSearchCV
from sklearn.svm import SVC
from sklearn.datasets import load_iris
from sklearn.decomposition import PCA
from sklearn.feature_selection import SelectKBest

iris = load_iris()

X, y = iris.data, iris.target

# This dataset is way too high-dimensional. Better do PCA:
pca = PCA(n_components=2)

# Maybe some original features where good, too?
selection = SelectKBest(k=1)

# Build estimator from PCA and Univariate selection:
combined_features = FeatureUnion([("pca", pca), ("univ_select", selection)])

# Use combined features to transform dataset:
X_features = combined_features.fit(X, y).transform(X)


svm = SVC(kernel="linear")

# Do grid search over k, n_components and C:
pipeline = Pipeline([("features", combined_features), ("svm", svm)])

param_grid = dict(features__pca__n_components=[1, 2, 3],
                  features__univ_select__k=[1, 2],
                  svm__C=[0.1, 1, 10])

grid_search = GridSearchCV(pipeline, param_grid=param_grid, cv=5, verbose=10)
grid_search.fit(X, y)
```

### mlr3pipelines

```{r, eval = FALSE}
library(mlr3)
library(mlr3pipelines)
library(mlr3learners)
library(mlr3featsel)
library(mlr3tuning)

iris = mlr_tasks$get("iris")

# Build the steps
copy = PipeOpCopy$new(2)
pca = PipeOpPCA$new()
selection = PipeOpFilter$new(filter = FilterVariance$new())
union = PipeOpFeatureUnion$new(2)
svm = PipeOpLearner$new(mlr_learners$get("classif.svm"))
svm$learner$param_set$values$kernel = "linear"

# Assemble the Pipeline
pipeline = copy %>>% gunion(list(pca, selection)) %>>% union %>>% svm
learner = GraphLearner$new(id = "Pipeline", pipeline)

# For tuning, we define the resampling and the Parameter Space
resampling = mlr3::mlr_resamplings$get("cv")
resampling$param_set$values$folds = 5
param_set = paradox::ParamSet$new(
  params = list(
    paradox::ParamDbl$new("classif.svm.cost", lower = 0.1, upper = 1),
    paradox::ParamInt$new("pca.rank.",  lower = 1, upper = 3),
    paradox::ParamInt$new("variance.filter.nfeat",  lower = 1, upper = 2)
  )
)

pe = PerformanceEvaluator$new(iris, learner, resampling, param_set)
terminator = TerminatorEvaluations$new(60)
tuner = TunerGridSearch$new(pe, terminator, resolution = 10)$tune()

# Set the learner to the optimal values and train
learner$param_set$values = tuner$tune_result()$values

e = Experiment$new(task = iris, learner = learner)
e$train()
```

In summary, we can achieve similar results with a comparable number of lines, while at the
same time offering greater flexibility with respect to which kinds of pipelines we want to
optimize over. At the same time, experiments using `mlr3` can now be arbitrarily parallelized
using `futures`.
