---
title: "Showcase: A simple pipeline"
author: "Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


This vignette showcases the general syntax and semantic behind `mlr3pipelines`.

## An introduction to PipeOperators

In this example, we create a ĺinear Pipeline. 
After scaling all input features, we rotate our data using principal component analysis.
After this transformation, we use a simple Decision Tree learner for classification.

As exemplary data, we will use the `iris` task.
This object contains the famous iris dataset and some meta-information, such as the target variable.

```{r}
  library(mlr3)
  task = mlr_tasks$get("iris")
```

We quickly split our data into a train and a test set: 
```{r}
  test.idx = sample(seq_len(task$nrow), 30)
  train.idx = setdiff(seq_len(task$nrow), test.idx)
  # Set task to only use train indexes
  task$row_roles$use = train.idx
```

A Pipeline (or `Graph`) contains multiple PipeOperators, where each `PipeOp`
transforms the data when it flows through it. For this usecase, we require 3 transformations:
- A PipeOp that scales the data
- A PipeOp that performs PCA
- A PipeOp that contains the **Decision Tree** learner

A list of available PipeOps can be obtained from

```{r}
  library(mlr3pipelines)
  mlr_pipeops$ids()
```

First we define the required PipeOps:

```{r}
  op1 = PipeOpScale$new()
  op2 = PipeOpPCA$new()
  op3 = PipeOpLearner$new(learner = mlr_learners$get("classif.rpart"))
```

### A quick glance into a PipeOp

In order to get a better understanding of what the respective PipeOps do, 
we quickly look at one of them in detail:

The following slots (and more) are thus contained in each PipeOp:
- `train`: A function used to train the PipeOp.
- `predict`: A function used to predict with the PipeOp.
- `id`: Set/Get the id of the PipeOp.
- `param_set`: The set of all exposed parameters of the PipeOp.
- `param_vals`: Current hyperparameter settings.
- `is_trained`: Is the PipeOp currently trained?

We can check the properties by accessing the respective slot.

```{r}
  op3$id
  op3$is_trained
```

The `param_set` and `param_vals` are required if a PipeOp contains
hyperparameters we want to set. See [ParamSet] for a quick intro on how `ParamSet`s work.

The `train()` and `predict()` functions define the core functionality of
our PipeOp. 
In many cases, in order to not leak information from the test set into the training set it is imperative to treat train and test data separately. For this we require a `train` function that learns the appropriate transformations from the training set and a `test` function that applies the transformation on future data.

In the case of `PipeOpLearner` this means the following:
- `train()` trains a model on its input  Task and saves the trained model to
  an additional slot, `.$state`. It returns a `list(NULL)`, as subsequent 
  operators usually do not require any output.
- `predict()` uses the model stored in `.$state` in order to predict
  the class of a new input task. It returns a [Prediction] object.
  This object contains the learner's predictions.

## Constructing the Pipeline

We can now connect the `PipeOp`s constructed earlier to a **Pipeline**.
We can do this using the `%>>%` operator.

```{r}
  linear_pipeline = op1 %>>% op2 %>>% op3
```

The result of this operation is a `graph`. 
A `graph` connects the input and output of each `PipeOp` to the following `PipeOp`.
This allows us to specify linear processing pipelines.
In this case, we connect the output of the **scaling** PipeOp to the input of the **PCA** PipeOp
and the output of the **PCA** PipeOp to the input of **PipeOpLearner**.

We can now train the `graph` using the iris **Task**.

```{r}
  linear_pipeline$train(task)
```

When we now train the graph, the data flows through the graph as follows: 
- The Task flows into the `PipeOpScale`. The `PipeOp` scales each column in the data
  contained in the Task and returns a new Task that contains the scaled data to its output.
  Relevant information

- The scaled task flows into the `PipeOpPCA`. PCA transforms the data and returns a (possibly smaller) 
  Task, that contains the transformed data. 

- This transformed data then flows into the learner, in our case **classif.rpart**.
  It is then used to train the learner, and as a result saves a model that can be 
  used to predict new data.

In order to predict on new data, we need to save the relevant transformations our data went through
while training. As a result, each PipeOp saves a state, where information requried to appropriately 
transform future data is stored. In our case, this is **mean** and **standard deviation** of each column for `PipeOpScale`, the PCA rotation matrix for `PipeOpPCA` and the learned model for `PipeOpLearner`.


```{r}
  # predict on test.idx
  task$row_roles$use = test.idx
  linear_pipeline$predict(task)
```
 
## Using the Pipeline as an mlr3 learner

In most cases, we want to use the pipeline just like an 
**mlr3** learner. In order to achieve this, 
we simply construct a `GraphLearner` that contains the
pipeline we defined.

```{r}
  graph_lrn = GraphLearner$new(linear_pipeline)
```

We can now use this learner in order to create an [mlr3::Experiment].

```{r}
  task = mlr_tasks$get("iris")
  e = Experiment$new(task = task, learner = graph_lrn)
```

For training:

```{r}
  e$train(train.idx)
  e$model
```

testing:

```{r}
  e$predict(test.idx)
  e$prediction
```

and scoring: 

```{r}
  e$score()
  e$performance["mmce"]
```

This allows us to seamlessly integrate pipelines with the
whole [mlr3] ecosystem, and thus for example resample or benchmark pipelines, tune its params and many more things.

For additional information on `mlr3` please refer to the respective vignettes.