---
title: "Showcase: Bagging and Stacking"
author: "Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>"
)
set.seed(8008135)
```

Leveraging the different operations available to connect `PipeOps` to very powerfull graphs.
This vignette introduces two such well-known graph structures, that allow us to enhance single learners
to more powerfull combinations.

Before we go into details, we split the task into train
and test indices.

```{r}
  library(mlr3)
  library(mlr3pipelines)
  task = mlr_tasks$get("iris")
  train.idx = sample(seq_len(task$nrow), 120)
  test.idx = setdiff(seq_len(task$nrow), train.idx)
```

## Bagging

We first examine Bagging introduced by Breimann (1994).
The basic idea is to create multiple predictors and then aggregate those to a single, more powerfull predictor.
> "... multiple versions are formed
> by making bootstrap replicates of the learning set
> and using these as new learning sets" (Breimann 1994)

Bagging then aggregates a set of predictors by averaging (regression) or majority vote (classification).
The idea behind bagging is, that a set of weak, but different predictors can be combined in order to arrive
at a single, better predictor.

We can achieve this by downsampling our data before training a learner, repeating this for say $10$ times and then performing a majority vote on the predictions.


First, we create a simple pipeline, that uses
`PipeOpSubsample` before a `PipeOpLearner` is trained:

```{r}
  single_pred = PipeOpSubsample$new() %>>%
    PipeOpLearner$new(mlr_learners$get("classif.rpart"))
```

We can now repeat this $10$ times using `greplicate`.

```{r}
  pred_set = greplicate(single_pred, 10L)
```

In order to visually inspect the resulting graph,
we can plot it:

```{r, fig.width=7.5}
  pred_set$plot()
```

Afterwards we need to aggregate the 10 pipelines
to form a single model:

```{r}
  bagging = pred_set %>>%
    PipeOpMajorityVote$new(innum = 10L)
```

and plot again to see what happens:

```{r, fig.width=7.5}
  bagging$plot()
```

This pipeline can again be used in conjunction with
`GraphLearner` in order for Bagging to be used like a
[mlr3::Learner].

```{r}
  baglrn = GraphLearner$new(bagging)
  e = Experiment$new(task = task, learner = baglrn)
  e$train(train.idx)
  e$predict(test.idx)
```

In conjunction with different `Backends`, this can be a very powerful tool, as in cases, where the data does not fully fit in memory, we can easily just obtain a fraction of the data for each learner from a `DataBaseBackend` and then aggregate predictions from all learners.

## Stacking

Stacking is another technique that can improve model performance. The basic idea behind stacking is, that
using predictions from one model as features of a subsequent model can possibly improve performance.

A very simple possibility would be to train a
decision tree and use the predictions from this
model in conjunction with the original features in order to train an additional model on top.
The basic idea behind this is, that patterns a model detected in the data can be used by a higher level model, and thus result in a better performance.

In order to limit overfitting, we additionally do not predict on
the original predictions of the learner, but instead on out-of-bag predictions. This is automatically done by
`PipeOpLearnerCV`.

We first create a level 0 learner, which is used to
extract a lower level prediction.
We additionally `clone()` the learner object to obtain a copy
of the learner, and set a custom id for the `PipeOp`.

```{r}
  lrn = mlr_learners$get("classif.rpart")
  lrn_0 = PipeOpLearnerCV$new(lrn$clone())
  lrn_0$id = "rpart_cv"
```

Additionally, we use a `PipeOpNULL` in parallel to
the level 0 learner, in order to send the unchanged
Task to the next level.

```{r}
  level_0 = gunion(list(lrn_0, PipeOpNULL$new()))
```

Afterwards, we want to concatenate the predictions from
`PipeOpLearnerCV` and the original Task using `PipeOpFeatureUnion`.

```{r}
  combined = level_0 %>>% PipeOpFeatureUnion$new(2)
```

We can now train another learner on top of the combined
features.

```{r, fig.width=7.5}
  stack = combined %>>% PipeOpLearner$new(lrn$clone())
  stack$plot()
```

```{r}
  stacklrn = GraphLearner$new(stack)
  e = Experiment$new(task = task, learner = stacklrn)
  e$train(train.idx)
  e$predict(test.idx)
```

In this vignette, we only showcase a very simple usecase for stacking.
In many real-world applications, stacking is done for multiple levels and on multiple representations of the dataset. On a lower level, different preprocessing methods can for example
be defined in conjunction with several learners. On a higher level, we can then combine
those predictions in order to form a very powerfull model.

## Multilevel Stacking

In order to showcase the power of `mlr3pipelines`, we will quickly introduce a more complicated
stacking example.

In this case, we train a **glmnet** and 2 different *rpart* models (some transform its inputs using `PipeOpPCA`) on our task in the level 0 and concatenate them with the original features (via `PipeOpNull`).
This is then passed on to level 1, where we copy the concatenated features
$3$ times and put this task into a **rpart** and a **glmnet** model.
Additionally, we also keep a version of the level 0 output (via `PipeOpNull`) and pass this on to level 2. In level 2 we simply concatenate all level 1 outputs and train a final decision tree.


```{r, eval = FALSE}
  library(mlr3learners)
  rprt = mlr_learners$get("classif.rpart")
  rprt$predict_type = "prob"
  glmn = mlr_learners$get("classif.glmnet")
  glmn$predict_type = "prob"

  #  Create Learner CV Operators
  lrn_0 = PipeOpLearnerCV$new(rprt, id = "rpart_cv_1")
  lrn_0$values$maxdepth = 5L
  lrn_1 = PipeOpPCA$new(id = "pca1") %>>% PipeOpLearnerCV$new(rprt, id = "rpart_cv_2")
  lrn_1$values$maxdepth = 1L
  lrn_2 = PipeOpPCA$new(id = "pca2") %>>% PipeOpLearnerCV$new(glmn)

  # Union them with a PipeOpNULL to keep original features
  level_0 = gunion(list(lrn_0, lrn_1,lrn_2, PipeOpNULL$new(id = "NULL1")))

  # Cbind the output 3 times, train 2 learners but also keep level
  # 0 predictions
  level_1 = level_0 %>>%
    PipeOpFeatureUnion$new(4) %>>%
    PipeOpCopy$new(3) %>>%
    gunion(list(
      PipeOpLearnerCV$new(rprt, id = "rpart_cv_l1"),
      PipeOpLearnerCV$new(glmn, id = "glmnt_cv_l1"),
      PipeOpNULL$new(id = "NULL_l1")
    ))

  # Cbind predicitions, train a final learner.
  level_2 = level_1 %>>%
    PipeOpFeatureUnion$new(3, id = "u2") %>>%
    PipeOpLearner$new(rprt,
      id = "rpart_l2")

  # Plot the resulting graph
  level_2$plot()

  e = Experiment$new(
    task = mlr_tasks$get("iris"),
    learner = GraphLearner$new(level_2)
    )
  e$train(train.idx)
  e$predict(test.idx)
  e$score()
  e$performance
```
